02_DESeq_characterization

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
new icon labeles plots

@@

  DEGs <- data %>% filter(direction != "NS")

@@

  volcano <- data %>%
    ggplot(aes(x = lfc, y = logpadj)) + 
    geom_point(aes(color = direction), size = 1, alpha = 0.75, na.rm = T) + 
     theme_B3() +
    scale_color_manual(values = mycolors,
                       name = " ",
                       drop = FALSE) +
    ylim(c(0,25)) +  
    xlim(c(-8,8)) +
    labs(y = "-log10(p)", x = "log fold change")  +
    theme(legend.position = c(.5, .95),
          legend.direction = "horizontal",
          legend.spacing.x = unit(-0.1, 'cm'),
          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
          panel.grid = element_blank()) 
  return(volcano)
}

@@

plot_grid(a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e ,
          f + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 2, rel_widths = c(0.25,0.2,0.2,0.2), rel_heights = c(0.45,0.55))

@@

>>>>

library(png)
library(grid)
library(ggimage)
library(ggpubr)

@@

  DEGs <- data %>% filter(direction != "NS")
  print(str(DEGs))

@@

  volcano <- data %>%
    ggplot(aes(x = lfc, y = logpadj)) + 
    geom_point(aes(color = direction), size = 1, alpha = 0.75, na.rm = T) + 
    theme_minimal() +
    scale_color_manual(values = mycolors,
                       name = " ",
                       drop = FALSE) +
    ylim(c(0,30)) +  
    xlim(c(-8,8)) +
    labs(y = "-log10 (p-value)", x = " ")  +
    theme(legend.position = "none",
          legend.direction = "horizontal",
          legend.spacing.x = unit(-0.1, 'cm'),
          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
          panel.grid = element_blank()) 
  return(volcano)
}

@@

females <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "female pituitary \n -log10 (p-value)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
females 
```

@@

## males and females

```{r createDEGsmales}
control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control")
bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg")
lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay") 
inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3") 
inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17") 
hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch") 
n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5")

```{r plotDEGsmales}
a2 <- plot.volcano(control.bldg2, "treatment", "bldg", "control", colorsvolcano)
b2 <- plot.volcano(bldg.lay2, "treatment", "lay", "bldg", colorsvolcano)
c2 <- plot.volcano(lay.inc.d32, "treatment", "inc.d3", "lay", colorsvolcano) 
d2 <- plot.volcano(inc.d3.inc.d92, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e2 <- plot.volcano(inc.d9.inc.d172, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f2 <- plot.volcano(inc.d17.hatch2, "treatment", "hatch", "inc.d17", colorsvolcano) 
g2 <- plot.volcano(hatch.n52, "treatment", "n5", "hatch", colorsvolcano) 
h2 <- plot.volcano(n5.n92, "treatment", "n9", "n5", colorsvolcano)

males <- plot_grid(a2 + labs(y = "male pituitary \n -log10 (p-value)"),
          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
males

femalesNmales <- plot_grid(females,males, nrow = 2, 
          rel_heights = c(0.45,0.55))

background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
background

icons <- ggdraw() +  draw_image(background, scale = 1)

plot_grid(icons, females, males, nrow = 3, rel_heights = c(1,1,1.5))

### tissue levels, ignoring sex differences, sort of

```{r gonaddfs, eval=T}

dds.gonads <- subsetDESeq2(c.colData,  c.countData, c("female_gonad","male_gonad"))

control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control")
bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg")
lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay") 
inc.d3.inc.d9 <- createDEGdf(dds.gonads) 
inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17") 
hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch") 
n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5") 
```

```{r gonplot, eval=T}
a2 <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b2 <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c2 <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d2 <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e2 <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f2 <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g2 <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h2 <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

gonads <- plot_grid(a2 + labs(y = "gonads \n -log10 (p-value)"),
          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
            labs(x = "log fold change"), 
          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
gonads
```

```{r hypdfs, eval=T}
dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus"))

control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control")
bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg")
lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay") 
inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus) 
inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17") 
hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch") 
n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5") 
```

```{r hypplot, eval = T}
a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

hypothalamus <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "hypothalamus \n -log10 (p-value)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
hypothalamus

plot_grid(nrow = 2)

plot_grid(icons, females, males, hypothalamus, gonads, nrow = 5, rel_heights = c(1,1,1,1,1.5))
